---
title: "Bike sharing"
output: html_notebook
---
#data prepare
```{r}
day <- read.csv("day.csv")
str(day)
```

```{r}
boxplot(day$cnt,main="total daily count")
```




the first column, instant, is just the list of case number, can be dropped
season,yr,month,weekday defines the dteday, so we do not want to include them

```{r}
cnt <- day[ , c(-1,-2,-4,-5,-7)]
str(cnt)
```

combine holiday and workingday cuz we want to analyze holiday.weekend as a whole
```{r}
cnt$holiday.weekend <- ifelse(cnt$holiday ==1 | cnt$workingday ==0, 1, 0)
str(cnt)
```

to make the codes easier,
```{r}
cnt <- cnt[ , c(-2,-3,-6)]
str(cnt)
```


#scatterplot

Scatterplot matrix to check colinearity
```{r}
library(psych)
pairs.panels(cnt[c("season","holiday.weekend","weathersit","temp","hum","windspeed","cnt")])
```
from the graph above
1. holiday.weekend may not be a significant predictor
2. temp and hum might be second order

check second order of temp
```{r}
scatter.smooth(cnt$temp,cnt$cnt,col="red")
```

check second order of hum
```{r}
scatter.smooth(cnt$hum,cnt$cnt,col="blue")
```



#AIC
```{r}
model_null <- lm(cnt~1,data=cnt)
model_full <- lm(cnt~factor(holiday.weekend)+factor(season)+factor(weathersit)+poly(temp,2, raw=FALSE)+poly(hum, 2, raw = FALSE)+windspeed,data=cnt)
```


```{r}
step(model_null,scope = list(lower=model_null,upper=model_full),direction = "forward")
```

#additive model
```{r}
model_additive<-lm(cnt~factor(season)+factor(holiday.weekend)+factor(weathersit)+temp+hum+windspeed,data=cnt)
summary(model_additive)
```

turns out holiday.weekend is not significant as guessed before
#drop the non-significant predictor
```{r}
model_additive2 <- lm(cnt~factor(season)+factor(weathersit)+temp+hum+windspeed,data=cnt)
summary(model_additive2)
```
try temp and season interaction
```{r}
model_interact <- lm(cnt~factor(season)*temp+factor(weathersit)+hum+windspeed,data=cnt)
summary(model_interact)
```


#polynomial model
```{r}
model_poly1 <- lm(cnt~poly(temp,2,raw = FALSE)+factor(season)+factor(weathersit)+hum+windspeed ,data=cnt)
summary(model_poly1)
```
```{r}
model_poly2 <- lm(cnt~poly(temp,2,raw = FALSE)+poly(hum,2,raw = FALSE)+factor(season)+factor(weathersit)+windspeed,data=cnt)
summary(model_poly2)
```

so final model is the two-quadratic model
```{r}
model_final <- model_poly2
summary(model_final)
```
```{r}
anova(model_final,model_additive2)
```
quadratic term is significant, do not drop


how could holiday.weekend not significant?
#compare casual and register Vs holiday.weekend
```{r}
par(mfrow=c(1,2))
boxplot(casual~holiday.weekend,data=cnt, main="casual users",xlab="holiday/weekend",ylab = "count")
boxplot(registered~holiday.weekend,data=cnt,main="registered users",xlab="holiday/weekend", ylab = "count")
```



#casual model
```{r}
library(psych)
pairs.panels(cnt[c("season","holiday.weekend","weathersit","temp","hum","windspeed","casual")])
```

```{r}
model_casual <- lm(casual~poly(temp,2,raw = FALSE)+poly(hum,2,raw = FALSE)+factor(season)+factor(weathersit)+windspeed+holiday.weekend,data=cnt)
summary(model_casual)
```

#registered
```{r}
library(psych)
pairs.panels(cnt[c("season","holiday.weekend","weathersit","temp","hum","windspeed","registered")])
```

```{r}
model_register <- lm(registered~poly(temp,2,raw = FALSE)+poly(hum,2,raw = FALSE)+factor(season)+factor(weathersit)+windspeed+holiday.weekend,data=cnt)
summary(model_register)
```

#diagnostics
```{r}
par(mfrow=c(2,2))
plot(model_final)
```

```{r}
shapiro.test(model_final$residuals)
```
```{r}
library(car)
ncvTest(model_final)
```


```{r}
library(car)
residualPlots(model_final)
```

```{r}
outlierTest(model_final)
```

```{r}
influencePlot(model_final)
```
delete 69 to see if the plots look better
```{r}
cnt_sub <- cnt[-69, ]
```

```{r}
model_sub <- lm(cnt~poly(temp,2,raw = FALSE)+poly(hum,2,raw = FALSE)+factor(season)+factor(weathersit)+windspeed,data=cnt_sub)
summary(model_sub)
```

```{r}
par(mfrow=c(2,2))
plot(model_sub)
```
```{r}
library(car)
ncvTest(model_sub)
```


